### 2012-08-02 Doc/VRIPHYS2012/vriphys2012.tex
small corrections
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@451 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
% TODO: anonymize -- Submission ID
\author[T. Golembiovský \& C. Duriez]
       {Tomáš Golembiovský$^{1,2}$ and Christian Duriez$^{2}$\\
        $^1$Faculty of Informatics, Masaryk University, Czech Republic\\
        $^2$INRIA Lille, France
       } % TODO: ask Christian how to specify affiliation

@@

\begin{abstract} % TODO
There is a strong need, in surgical simulations, for physics-based deformable model of thin or hollow structures.
The use of shell theory allows to have a well-founded formulation resulting from continuum mechanics of thin objects.
However, this formulation asks for second order spatial derivatives so requires the use of complex elements.
In this paper, we present a new way of building the interpolation:
First, we use the cubic B/'ezier shell triangle to allow for a good continuity inside and between the elements and second, we build a kinematic mapping to reduce the degrees of freedom of the element from 10 control points with 3 Degrees of Freedom ($=30$ DoFs) to only 3 nodes with 6 DoFs ($=18$ DoFs).
This reduction allows for good computation performance.
This new shell model description is also used to map a smooth surface (for the collision detection and response) on a coarse mechanical mesh to accounts for the complex contacts that take place during surgical procedures.
We demonstrate the convergence and the computational efficiency of our approach as well as its use in two different simulation cases: the planning of surgery for congenital heart disease correction and a preliminary simulation of childbirth.

\begin{classification} % according to http://www.acm.org/class/1998/
    \CCScat{Computer Graphics}{I.3.5}{Computational Geometry and Object Modeling}{Physically based modelling};
    \CCScat{Computer Graphics}{I.3.7}{Three-Dimensional Graphics and Realism}{Animation}.
    %\CCScat{Simulation and Modeling}{I.6.8}{Types of Simulation}{Animation}
    %\CCScat{Life and Medical Sciences}{J.3}{Health}{}
\end{classification}

@@

This problem has been faced by the computer graphics community but for different applications.
For instance, many related works address the problem of the simulation of cloth using mass-spring models \cite{Volino2009} or bending models \cite{Grinspun2003}. %% Todo Choi2007?
However, these models are discrete so their behavior depends on the mesh and their parameters are not easily measurable.
This is completely acceptable for some computer graphics application, where the parameters can be tuned manually to obtain a realistic animation.
In the case of medical simulation, more physics-based modeling is needed.

@@

\subsection{Projection of surface forces}

@@

Second performed test is an hemisphere with hole subjected to two opposing
forces at the base. The geometry is described by a hemisphere with radius
$r=10$ and thickness $t=0.04$ with 18° hole at the top and the geometry is
discretized into the grid of NxN vertices per quadrant. The physical
parameters are $E = 6.825 \times 10^7$ and $\nu = 0.3$. The hemisphere is
at it's base subjected to two opposing outwards forces and two opposing
inwards forces with magnitude $P = 4$. To avoid rigid body movement we have
constrained two opposite nodes at the top of the hemisphere. The radial
displacement for one point with applied load is shown in Figure
\ref{fig-hemisphere}.

@@

% TODO?
%Another designed application is simulation of women reproductive organs
%(vagaina/uterus/\ldots). 
%The simulation is so as to evaluate
%stress on a tissue during a birth.

\subsection{Simulation of uterus deformation during birth}

@@

Our study relies on the assumption that the in-plane and the bending deformation energies can be separated.
For complex deformation fields or material constitutive laws  this assumption is no more valid.
However, we could keep the same way of building the element even with more complex mechanics. 
For instance the interpolation would allow for derivation of three-dimensional non-linear strain tensors.
We will investigate this topic in our future work.
% }}}

@@

In this paper we presents new shell element with interpolation functions
based on B\'ezier triangle which is  suitable for real-time simulations.
Compared to  the elements commonly used in mechanics, the interpolation is less complex and ensure at least a $C^0$ continuity in the worst case. 
(Many shell elements are not even $C^0$ when the initial shape is not flat).

@@

Because of the formulation of B\'ezier triangle is intrinsically present
there is a clean and obvious way of mapping high-resolution mesh for better
visualization. We also provide a technique of computing velocities of any point
on the surface and formula for back-projection of forces applied on the
surface onto the mechanical mesh. 
Those are essential components for the surgical simulators that need accurate contact geometry modeling without an overflow of computation need
The efficiency of the approach is demonstrated through convergence tests,  computation time benchmark and application to two different  simulation cases.

>>>>

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
% TODO: anonymize -- Submission ID
%\author[paper1026]
%       {paper1026}
\author[T. Golembiovský \& C. Duriez]
       {Tomáš Golembiovský$^{1,2}$ and Christian Duriez$^{2}$\\
        $^1$Faculty of Informatics, Masaryk University, Czech Republic\\
        $^2$INRIA Lille, France
       } % TODO: ask Christian how to specify affiliation

@@

\begin{abstract}
There is a strong need, in surgical simulations, for physically based deformable model of thin or hollow structures.
The use of shell theory allows to have a well-founded formulation resulting from continuum mechanics of thin objects.
However, this formulation asks for second order spatial derivatives so requires the use of complex elements.
In this paper, we present a new way of building the interpolation:
First, we use the trianular cubic B\'ezier shell to allow for a good continuity inside and between the elements and second, we build a kinematic mapping to reduce the degrees of freedom of the element from 10 control points with 3 Degrees of Freedom ($=30$ DOFs) to only 3 nodes with 6 DOFs ($=18$ DOFs).
This reduction allows for good computation performance.
This new shell model description is also used to map a smooth surface (for the collision detection and response) on a coarse mechanical mesh to account for the complex contacts that take place during surgical procedures.
We demonstrate the convergence and the computational efficiency of our approach as well as its use in two different simulation cases: the planning of surgery for congenital heart disease correction and a preliminary simulation of childbirth.

\begin{classification} % according to http://www.acm.org/class/1998/
    \CCScat{Computer Graphics}{I.3.5}{Computational Geometry and Object Modeling}{Physically based modelling}.
    %\CCScat{Computer Graphics}{I.3.7}{Three-Dimensional Graphics and Realism}{Animation}.
    %\CCScat{Simulation and Modeling}{I.6.8}{Types of Simulation}{Animation}
    %\CCScat{Life and Medical Sciences}{J.3}{Health}{}
\end{classification}

@@

This problem has been faced by the computer graphics community but for different applications.
For instance, many related works address the problem of the simulation of cloth using mass-spring models \cite{Volino2009} or bending models \cite{Grinspun2003}. %% Todo Choi2007?
However, these models are discrete so their behavior depends on the mesh and their parameters are not easily measurable.
This is completely acceptable for some computer graphics application, where the parameters can be tuned manually to obtain a realistic animation.
In the case of medical simulation, more physically based modeling is needed.

@@

\subsection{Projection of Surface Forces}

@@

Second performed test is a hemisphere with hole subjected to two opposing
forces at the base. The geometry is described by a hemisphere with radius
$r=10$ and thickness $t=0.04$ with 18° hole at the top and the geometry is
discretized into the grid of NxN vertices per quadrant. The physical
parameters are $E = 6.825 \times 10^7$ and $\nu = 0.3$. The hemisphere is
at it's base subjected to two opposing outwards forces and two opposing
inwards forces with magnitude $P = 4$. To avoid rigid body movement we have
constrained two opposite nodes at the top of the hemisphere. The radial
displacement for one point with applied load is shown in Figure
\ref{fig-hemisphere}.

@@

\subsection{Simulation of Uterus Deformation During Birth}

@@

Our study relies on the assumption that the in-plane and the bending deformation energies can be separated.
For complex deformation fields or material constitutive laws  this assumption is no more valid.
However, we could keep the same way of building the element even with more complex mechanics. 
For instance the interpolation would allow for derivation of three-dimensional non-linear strain tensors.
We will investigate this topic in our future work.

@@

In this paper we presents new shell element with interpolation functions
based on B\'ezier triangle which is  suitable for real-time simulations.
Compared to  the elements commonly used in mechanics, the interpolation is less complex and ensures at least a $C^0$ continuity in the worst case. 
(Many shell elements are not even $C^0$ when the initial shape is not flat).

@@

Because of the formulation of B\'ezier triangle is intrinsically present
there is a clean and obvious way of mapping high-resolution mesh for better
visualization. We also provide a technique of computing velocities of any point
on the surface and formula for back-projection of forces applied on the
surface onto the mechanical mesh. 
Those are essential components for the surgical simulators that need accurate contact geometry modeling without an overflow of computation.
The efficiency of the approach is demonstrated through convergence tests,  computation time benchmark and application to two different  simulation cases.
